Effects

### 2019-12-30 doc/Effects.Rmd
working on Effect doc

load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")
#save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")

@@

## Quality Effects

@@

#ad
#  geom_smooth(method = lm, se = FALSE
 # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)

@@

I'm not sure sure how effective the above plots are.

### Age Effect (male)

@@

@@

### Age Effect (female)

@@

### Maternal Age effect

@@

@@

@@

Compare observations from the same mouse, but different MLH1 stains.

```{r stain.effects, eval=FALSE}

#compile this dataset

>>>>

load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")

#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")
#save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")

@@

# Quality Effects

@@

#ad
#  geom_smooth(method = lm, se = FALSE
 # geom_smooth(method = lm, formula = mean.MLH1.obs$mean.MLH1 ~ mean.MLH1.obs$category, se = FALSE)
```

```{r qual.tables, echo=FALSE}

#make a similart plot as the ages
#for each mouse and quality, calq the mean MLH1
mouse.table.mean.MLH1.obs_qual <- ddply(.data=MLH1_data, 
                 .(mouse, quality),
                 summarize, 
                 ncells =  length(unique(Original.Name)),
                 mean.MLH1 = mean(adj_nMLH1.foci),
                 var.MLH1 = var(adj_nMLH1.foci)
)

mouse.table.mean.MLH1.obs_qual <- add_category(mouse.table.mean.MLH1.obs_qual)
mouse.table.mean.MLH1.obs_qual <- add_strain(mouse.table.mean.MLH1.obs_qual)
mouse.table.mean.MLH1.obs_qual <- add_sex(mouse.table.mean.MLH1.obs_qual)
mouse.table.mean.MLH1.obs_qual <- add_subsp(mouse.table.mean.MLH1.obs_qual)

@@

```{r qual.plot, echo=FALSE}

mm <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "male",]
age.effect.plot.mm <- ggplot(mm, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)
age.effect.plot.mm

ff <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "female",]
age.effect.plot.ff <- ggplot(ff, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female quality and mean CO")+facet_wrap(~subsp)

age.effect.plot.ff

The original plots weren't very effective, so I removed them. I made

## Quality effect on SC

```{r SC.qual, echo=FALSE}
#use the SC table

#all of skel and bin are not num
MLH1.merge.Skel$skel_size <- as.character(MLH1.merge.Skel$skel_size)
MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)

MLH1.merge.Skel$bin_size <- as.character(MLH1.merge.Skel$bin_size)
MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)

#mouse.level.skel

#MLH1.merge.Skel, hist of all skel values

#test with removing crop == 'no', MLH1.merge.Skel[MLH1.merge.Skel$REDO.crop == "no",]
mouse.table.mean.skel.obs_qual <- ddply(.data=MLH1.merge.Skel, 
                 .(mouse, quality),
                 summarize, 
                 ncells =  length(unique(Original.Name)),
                 mean.skel = mean(skel_size, na.rm = TRUE),
                 mean.bin = mean(bin_size)
                # var.skel = var(skel_size, na.rm = TRUE) #var not useful, many NA/1 obs
)

mouse.table.mean.skel.obs_qual <- add_category(mouse.table.mean.skel.obs_qual)
mouse.table.mean.skel.obs_qual <- add_strain(mouse.table.mean.skel.obs_qual)
mouse.table.mean.skel.obs_qual <- add_sex(mouse.table.mean.skel.obs_qual)
mouse.table.mean.skel.obs_qual <- add_subsp(mouse.table.mean.skel.obs_qual)

```{r SC.hitograms, echo=FALSE}

skel.histgram <- ggplot(MLH1.merge.Skel, aes(x=skel_size)) +  geom_histogram(bins=50)
skel.histgram

length(MLH1.merge.Skel$Random.Name[MLH1.merge.Skel$skel_size > 30000])#67 large SC cells

MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)

MLH1.merge.Skel <- MLH1.merge.Skel[MLH1.merge.Skel$skel_size < 30000,]

skel.histgrams.2.DOM <- ggplot(MLH1.merge.Skel[MLH1.merge.Skel$subsp == "Dom",], aes(x=skel_size)) +  geom_histogram(bins=50)+facet_grid(sex~strain)
skel.histgrams.2.DOM

skel.histgrams.2.MUSC <- ggplot(MLH1.merge.Skel[MLH1.merge.Skel$subsp == "Musc",], aes(x=skel_size)) +  geom_histogram(bins=50)+facet_grid(sex~strain)
skel.histgrams.2.MUSC

```{r sc.qual.plots, echo=FALSE}
skel.mm <- mouse.table.mean.skel.obs_qual[mouse.table.mean.skel.obs_qual$sex == "male",]
skel.mm <- skel.mm[skel.mm$mean.skel < 50000,]

skel.qual.effect.plot.mm <- ggplot(skel.mm, aes(y=mean.bin, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)
skel.qual.effect.plot.mm
```

Each point is a single mouse

```{r G.plot, echo=FALSE}
Bin.qual.effect.plot.mm.G <- ggplot(skel.mm[skel.mm$strain == "G",],
                           aes(y=mean.bin, x=quality, size= ncells, color=strain))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Quality and mean SC (bin) ")+facet_wrap(~mouse)

Bin.qual.effect.plot.mm.G

skel.qual.effect.plot.mm.G <- ggplot(skel.mm[skel.mm$strain == "G",],
                           aes(y=mean.skel, x=quality, size= ncells, color=strain))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Quality and mean SC (bin) ")+facet_wrap(~mouse)

The mouse 22jun15_G_m2 has strange pattern, investigate these outputs

Some of the pattern goes away when redo.crop is set to equal 'no'.

```{r female.skel.plot, echo=FALSE}

skel.ff <- mouse.table.mean.skel.obs_qual[mouse.table.mean.skel.obs_qual$sex == "female",]

skel.qual.effect.plot.ff <- ggplot(skel.ff, aes(y=mean.bin, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)

skel.qual.effect.plot.ff

skel.qual.effect.plot.ff.MSM <- ggplot(skel.ff[skel.ff$strain == "MSM",], aes(y=mean.bin, x=quality, color=(strain), size=ncells ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)

skel.qual.effect.plot.ff.G <- ggplot(skel.ff[skel.ff$strain == "G",], aes(y=mean.bin, x=quality, color=(strain), size=ncells ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)

skel.qual.effect.plot.ff.PWD <- ggplot(skel.ff[skel.ff$strain == "PWD",], aes(y=mean.skel, x=quality, color=(strain), size=ncells ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~mouse)

The above plots are good for detecting outliers

# Age Effect (male)

@@

ADD euthdate?

@@

# Age Effect (female)

@@

# Maternal Age effect

@@

mouse.table.female <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.x == "female",]

mouse.table.female <- mouse.table.female[!(is.na(mouse.table.female$mouse) | mouse.table.female$mouse==""), ]

mouse.table.female <- mouse.table.female[!(is.na(mouse.table.female$euth.date) | mouse.table.female$euth.date==""), ]

#add euth dates to MLH1_data
test.this <- MLH1_data[ MLH1_data$mouse %in% mouse.table.female$mouse, ]

mlh1.euth.date.female <- merge(mouse.table.female, MLH1_data, by = "mouse")

@@

```{r plot.dissection, echo=FALSE}

male.data$euth.date <- as.Date(male.data$euth.date, "%m/%d/%Y")

male.euth.date <- ggplot(male.data, aes(y=mean_co, x=euth.date, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)
male.euth.date

#look at musc
musc.male.euth.date <- ggplot(male.data[male.data$subsp.x == "Musc",], aes(y=mean_co, x=euth.date, color=(age.weeks) ))+geom_point()+ ylim(c(19,34))+
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Musc Male Age and mean CO")+facet_wrap(~strain.x)
musc.male.euth.date

```{r dissection.female, echo=FALSE}

female.data <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.y == "female") & (AP_mouse_table_w.Ages$category.y != "other"),]
female.data$euth.date <- as.Date(female.data$euth.date, "%m/%d/%Y")

female.euth.date.plot <- ggplot(female.data[female.data$subsp.x == "Musc",], aes(y=mean_co, x=euth.date, color=(strain.y) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male Age and mean CO")+facet_wrap(~subsp.y)

female.euth.date.plot

DOM.female.euth.date.plot <- ggplot(female.data[female.data$subsp.x == "Dom",], aes(y=mean_co, x=euth.date, color=(strain.y) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Age and mean CO")+facet_wrap(~subsp.y)

DOM.female.euth.date.plot

#female / male, sex.x
#mouse.table.male <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.x == "male",]

test.this <- MLH1_data[ MLH1_data$mouse %in% mouse.table.female$mouse, ]

mlh1.euth.date.female <- merge(mouse.table.female, MLH1_data, by = "mouse")

## Total SC

follow up to peromyscus findings, Where summer caught mice had lower SC area per cell, compared to winter and fall caught. But there were few data points.

```{r total.sc, echo=FALSE}
#check out effects of month dissected on SC

#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating

#from total SC

#merge
ll <- merge(mouse.level.skel, AP_mouse_table_w.Ages, by = "mouse")

ll$euth.date.2 <- as.Date(ll$euth.date, "%m/%d/%Y")

ll <- ll[!(is.na(ll$mean.skel) | ll$mean.skel==""), ]
ll <- ll[!(is.na(ll$sex.y) | ll$sex.y==""), ]

library(lubridate)

#ADD month 
ll$month.2 <- month(ll$euth.date.2)

#mouse level table

ll.plot <- ggplot(ll, aes(y=mean.skel, x=euth.date.2, color=(strain.y) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex.y)

ll.month.plot <- ggplot(ll, aes(y=mean.skel, x=month.2, color=(strain.y) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex.y)

ll.month.plot

```{r total.sc.all.cells, echo=FALSE, include=FALSE}
#try ploting all cells

#uu <- merge(MLH1.merge.Skel,AP_mouse_table_w.Ages,by="mouse"  )

#uu$euth.date.2 <- as.Date(uu$euth.date, "%m/%d/%Y")
#uu$month.2 <- month(uu$euth.date.2)

#uu <- uu[!(is.na(uu$skel_size) | uu$skel_size==""), ]
#uu <- uu[!(is.na(uu$sex.x) | uu$sex.x==""), ]

#uu.month.plot.DOM <- ggplot(uu[uu$subsp.x == "Dom",], aes(y=skel_size, x=month.2, color=(mouse) ))+
#  geom_jitter()+  theme(legend.position="none")+
#  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+
#  facet_wrap(~sex.x)

So, there doesn't seem to be an effect of month dissected on SC area across mice. This might be the start of testing for 'technical effects / humidity effects'. BUT, I don't really have month by month dissections of mice of the same age, ect.

@@

Compare observations from the same mouse, but different MLH1 stains.  First need to identify data that can be used.

```{r stain.effects.SC, eval=FALSE}
#/src/TotalSC.R
#compile this dataset
